---
title: "Replicating Chetty et al. (2014): Classical vs Permutation Inference for Intergenerational Mobility"
author: "Max Miller"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: false
    toc: true
    toc_depth: 2
header-includes:
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,breakanywhere,commandchars=\\\{\}}
  - \DefineVerbatimEnvironment{verbatim}{Verbatim}{breaklines,breakanywhere}
  - \usepackage{geometry}
  - \geometry{margin=0.75in}
  - \usepackage{listings}
  - \lstset{breaklines=true, breakatwhitespace=true, basicstyle=\ttfamily\small}
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 55),
  comment = NA
)
options(width = 70)
```

# Introduction

## Background: The Chetty et al. Study

In their landmark 2014 paper "Where is the Land of Opportunity? The Geography of Intergenerational Mobility in the United States," Chetty, Hendren, Kline, and Saez analyzed tax records from over 40 million children to measure intergenerational economic mobility across U.S. Commuting Zones (CZs).

A key finding was identifying which community-level characteristics correlate with **relative mobility**---the relationship between a child's income rank and their parents' income rank. Online Appendix Table VIII of their paper reports correlations between relative mobility and 35 covariates spanning categories like segregation, income inequality, education, family structure, and social capital.

## Purpose of This Analysis

This analysis has two objectives:

1. **Replicate** the classical OLS regression results from Chetty et al.'s Online Appendix Table VIII
2. **Compare** these classical results to a permutation-based inference method that does not rely on asymptotic normality assumptions

## Why Permutation Tests?

Classical OLS regression relies on asymptotic theory---as sample sizes grow large, the t-statistic approximately follows a t-distribution. However, this approximation may be unreliable when:

- Sample sizes are moderate (here, n $\approx$ 700 CZs)
- Error distributions are non-normal or heteroskedastic
- The relationship between variables violates model assumptions

**Permutation tests** offer an alternative that:

- Makes no distributional assumptions about the errors
- Provides exact p-values under the null hypothesis of no relationship
- Is robust to heteroskedasticity when properly studentized

We implement the permutation regression test from **DiCiccio and Romano (2017)**, "Robust Permutation Tests for Correlation and Regression Coefficients," which provides a heteroskedasticity-robust test statistic.

## The Data: Tables 5 and 8

Our analysis uses two of Chetty et al.'s Online Data Tables:

**Online Data Table 5** provides the **dependent variable**: relative mobility estimates for 741 Commuting Zones (CZs). These estimates were computed from IRS tax records covering over 40 million children born between 1980-1991, linked to their parents' tax returns. The key variable is the **rank-rank slope**---the regression coefficient when regressing children's income percentile rank on parents' income percentile rank within each CZ. A higher slope indicates less mobility (children's outcomes more strongly determined by parents' income). The national average is 0.34, meaning a 10 percentile increase in parent rank is associated with a 3.4 percentile increase in child rank. We use the 1980-82 birth cohort estimates and restrict to CZs with at least 250 children to ensure reliable estimates.

**Online Data Table 8** provides the **independent variables**: 35 community-level characteristics for each CZ, including measures of segregation, income inequality, school quality, social capital, family structure, and local labor market conditions. These covariates come from public sources (Census, IRS public statistics, education databases) and allow us to ask: "Which community characteristics predict higher or lower mobility?"

## Why Replicate Table VIII?

We chose to replicate Online Appendix Table VIII because it represents the core empirical contribution of Chetty et al. (2014): identifying which community-level characteristics are associated with intergenerational mobility. While the mobility estimates in Table 5 were computed from restricted IRS tax records and cannot be independently reproduced from raw data, Table VIII's correlational analysis uses these *published* mobility estimates as the dependent variable and publicly available community characteristics as predictors. This makes it both replicable and substantively important---it is this analysis that identifies single-parent household rates as "the single strongest correlate of upward income mobility among all the variables we explored" (Chetty et al., 2014, p. 40), and that links segregation, inequality, and social capital to children's economic outcomes. Policy discussions citing this paper typically draw on Table VIII's findings, making it essential to verify that these correlations are robust to alternative statistical methods.

## A Note on P-Values and Hypothesis Testing

A p-value greater than 0.05 **never** leads to rejecting the null hypothesis when using $\alpha = 0.05$ as the significance threshold. By definition, we reject $H_0$ only when $p < \alpha$. When $p \geq 0.05$, we "fail to reject" the null---this is not the same as accepting the null or proving no effect exists; it simply means the observed data are not sufficiently improbable under $H_0$ to warrant rejection at our chosen confidence level.

Some important clarifications:

- **Different $\alpha$ levels**: If a researcher uses $\alpha = 0.10$, then $p = 0.07$ would lead to rejection. The choice of $\alpha$ is made before seeing the data.
- **One-tailed vs. two-tailed tests**: A two-tailed $p = 0.08$ corresponds to a one-tailed $p = 0.04$, which would reject under a one-tailed test. However, our analysis uses two-tailed tests throughout.
- **Multiple testing**: When testing many hypotheses (as we do with 35 covariates), some researchers apply corrections like Bonferroni ($\alpha_{adjusted} = 0.05/35 \approx 0.0014$), which would make rejection even more stringent.

In this analysis, we use the conventional $\alpha = 0.05$ two-tailed threshold without multiple testing correction, consistent with Chetty et al.'s original analysis.

## Research Question

> Do the significance conclusions from Chetty et al.'s classical OLS analysis hold up under permutation-based inference?

If both methods agree, this provides evidence that the original findings are robust to the choice of inference method.

# Methods

## Data Sources

We use two publicly available datasets from the Equality of Opportunity Project:

| Dataset | Contents |
|---------|----------|
| Online Data Table 5 | Mobility estimates (relative and absolute) for each CZ |
| Online Data Table 8 | CZ-level characteristics (35 covariates) |

Following Chetty et al., we restrict the sample to CZs with at least 250 children (n $\geq$ 250) to ensure reliable mobility estimates.

## Statistical Methods

### Method 1: Classical OLS Regression (Chetty et al.'s Approach)

For each covariate $X_j$, we estimate:

$$\text{Relative Mobility}_i = \alpha + \beta_j X_{ji} + \epsilon_i$$

where both the outcome and covariate are standardized (mean 0, SD 1). The standardized coefficient $\beta_j$ equals the Pearson correlation coefficient $r$.

Statistical significance is assessed using the classical t-test, which assumes $\epsilon_i \sim N(0, \sigma^2)$.

### Method 2: Permutation Regression Test (DiCiccio & Romano, 2017)

The permutation test assesses significance by:

1. Computing a heteroskedasticity-robust Wald statistic $S_n$ from the observed data
2. Permuting the outcome variable $Y$ many times (B = 10,000) to generate a null distribution
3. Computing the p-value as the proportion of permuted statistics $\geq$ the observed statistic

The test statistic is:

$$S_n = n \hat{\beta}^T \left( \hat{\Sigma}_{XX}^{-1} \hat{\Omega} \hat{\Sigma}_{XX}^{-1} \right) \hat{\beta}$$

where $\hat{\Omega}$ is the heteroskedasticity-robust covariance estimator. This follows Theorem 3.2 of DiCiccio and Romano (2017).

**Key advantages:**

- No assumption of normally distributed errors
- Robust to heteroskedasticity
- Exact Type I error control under the null hypothesis

## Console Output Format

Results are printed in the following format:

```
[OK] Racial.Segregation  b=0.406 p_classical=0.000 p_perm=0.000*
```

| Symbol | Description |
|--------|-------------|
| `[OK]` | Classical and permutation methods agree on significance |
| `[X]` | Methods disagree on significance |
| `b` | Standardized regression coefficient (= Pearson r) |
| `p_classical` | p-value from classical OLS t-test |
| `p_perm` | p-value from permutation regression test |
| `*` | Statistically significant at $\alpha = 0.05$ |

# Part 1: Data Loading and Preparation

## Load Required Libraries

```{r libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
```

## Load the Chetty et al. Data

The data files are structured with metadata rows at the top, which we skip during import.

```{r load-data}
# Load Online Data Table 5 (Mobility Estimates)
# Row 50 contains column headers, so skip first 49 rows
table5 <- read.csv("Online_Data_Table_5.csv",
  skip = 49,
  header = TRUE,
  stringsAsFactors = FALSE
)

# Remove rows with column numbers "(1)", "(2)", etc. and empty rows
table5 <- table5[!grepl("^\\(\\d+\\)$", table5[, 1]) & table5[, 1] != "", ]
rownames(table5) <- NULL

# Load Online Data Table 8 (CZ Characteristics)
# Row 7 contains column headers, so skip first 6 rows
table8 <- read.csv("Online_Data_Table_8.csv",
  skip = 6,
  header = TRUE,
  stringsAsFactors = FALSE
)

# Remove invalid rows (negative CZ codes are metadata)
table8 <- table8[as.numeric(table8[, 1]) > 0 & !is.na(table8[, 1]), ]
rownames(table8) <- NULL

if (nrow(table8) == 0) {
  stop("ERROR: Online_Data_Table_8.csv loaded 0 rows! Check file location.")
}

cat(sprintf("Loaded Table 5: %d CZs with mobility estimates\n", nrow(table5)))
cat(sprintf("Loaded Table 8: %d CZs with characteristics\n", nrow(table8)))
```

## Clean and Prepare Data

```{r clean-data}
# Clean column names (R converts spaces and special chars to dots)
names(table5) <- gsub("^X\\.", "", names(table5))
names(table8) <- gsub("^X\\.", "", names(table8))

# Rename key columns for clarity
colnames(table5)[1] <- "CZ"
colnames(table5)[5] <- "AM"          # Absolute Mobility
colnames(table5)[6] <- "RM"          # Relative Mobility
colnames(table5)[4] <- "n_children"
colnames(table8)[1] <- "CZ"

# Convert CZ to numeric for joining
table5$CZ <- as.numeric(table5$CZ)
table8$CZ <- as.numeric(table8$CZ)
```

## Merge Tables and Apply Sample Restrictions

Following Chetty et al., we restrict to CZs with at least 250 children to ensure reliable mobility estimates.

```{r merge-tables}
# Merge mobility estimates with CZ characteristics
data <- table5 %>%
  select(CZ, RM, AM, n_children) %>%
  inner_join(table8, by = "CZ")

# Convert covariate columns to numeric
numeric_cols <- c("RM", "AM", "n_children", names(table8)[4:ncol(table8)])
for (col in numeric_cols) {
  if (col %in% names(data)) {
    data[[col]] <- as.numeric(data[[col]])
  }
}

# Apply sample restriction: n >= 250 children
data <- data %>%
  mutate(n250 = ifelse(n_children >= 250, 1, 0))

df <- data %>%
  filter(n250 == 1) %>%
  filter(!is.na(RM))

cat(sprintf("Final sample: %d CZs with n >= 250 children\n", nrow(df)))
cat(sprintf("(Dropped %d CZs with insufficient sample size)\n", nrow(data) - nrow(df)))
```

# Part 2: Permutation Test Implementation

## The DiCiccio & Romano (2017) Permutation Regression Test

This function implements Theorem 3.2 from DiCiccio and Romano (2017). The key innovation is using a **studentized test statistic** that is asymptotically pivotal, meaning the permutation distribution approximates the true null distribution even without exchangeability.

The null hypothesis is $H_0: \beta = 0$ (no linear relationship between X and Y).

```{r perm-functions}
#' Permutation Regression Test (DiCiccio & Romano 2017, Theorem 3.2)
#'
#' Tests H0: beta = 0 using a heteroskedasticity-robust Wald statistic.
#' The test permutes Y values to generate the null distribution.
#'
#' @param X Matrix of predictor variables (must be centered to mean zero)
#' @param Y Vector of response variable (must be centered to mean zero)
#' @param B Number of permutations (default: 10,000)
#' @param seed Random seed for reproducibility
#' @return List with observed statistic, permutation distribution, and p-value

perm_test_regression <- function(X, Y, B = 10000, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)

  X <- as.matrix(X)
  Y <- as.numeric(Y)
  n <- nrow(X)

  if (n != length(Y)) stop("X and Y must have same number of rows")

  # Always include an intercept (per Remark 3.1 of the paper)
  X <- cbind(1, X)

  # Compute the heteroskedasticity-robust Wald statistic S_n
  compute_Sn <- function(X, Y) {
    XtX <- crossprod(X)              # X'X
    XtY <- crossprod(X, Y)           # X'Y
    beta_hat <- solve(XtX, XtY)      # OLS estimate
    resid <- as.vector(Y - X %*% beta_hat)

    Sigma_hat <- XtX / n             # (1/n) X'X
    r2 <- resid^2
    Omega_hat <- crossprod(X, X * r2) / n  # HC0 robust covariance

    tmp <- solve(Sigma_hat, beta_hat)
    as.numeric(n * t(tmp) %*% Omega_hat %*% tmp)
  }

  # Observed test statistic
  S_obs <- compute_Sn(X, Y)

  # Generate permutation distribution under H0
  S_perm <- replicate(B, {
    Y_perm <- sample(Y, replace = FALSE)  # Permute Y values
    compute_Sn(X, Y_perm)
  })

  # Compute p-value (proportion of permuted stats >= observed)
  # +1 correction ensures valid p-value (Phipson & Smyth, 2010)
  pval <- (sum(S_perm >= S_obs) + 1) / (B + 1)

  list(S_obs = S_obs, S_perm = S_perm, p_value = pval, B = B)
}
```

# Part 3: Define Covariates from Table VIII

## The 35 Covariates

These are the same covariates analyzed in Online Appendix Table VIII of Chetty et al. (2014), organized by thematic category.

```{r covariates}
# Column names match R's processing of CSV headers
# (spaces -> dots, special characters -> dots)
covariates_to_test <- list(

  # RACE (1 variable)
  Race = c(
    "Frac..Black"                          # Fraction Black Residents
  ),

  # SEGREGATION (5 variables)
  Segregation = c(
    "Racial.Segregation",                  # Racial Segregation (Theil Index)
    "Income.Segregation",                  # Income Segregation (Theil Index)
    "Segregation.of.Poverty...p25.",       # Segregation of Poverty (<p25)
    "Segregation.of.Affluence...p75.",     # Segregation of Affluence (>p75)
    "Frac..with.Commute...15.Mins"         # Share with Commute < 15 Mins
  ),

  # INCOME DISTRIBUTION (5 variables)
  Income_Distribution = c(
    "Household.Income.per.capita",         # Household Income per Capita
    "Gini",                                # Gini Coefficient
    "Top.1..Income.Share",                 # Top 1% Income Share
    "Gini.Bottom.99.",                     # Gini Bottom 99%
    "Frac..Between.p25.and.p75"            # Fraction Middle Class
  ),

  # K-12 EDUCATION (4 variables)
  K12_Education = c(
    "School.Expenditure.per.Student",      # School Expenditure per Student
    "Teacher.Student.Ratio",               # Teacher-Student Ratio
    "Test.Score.Percentile..Income.Adjusted",    # Test Scores (Income Adjusted)
    "High.School.Dropout.Rate..Income.Adjusted." # Dropout Rate (Income Adjusted)
  ),

  # SOCIAL CAPITAL (3 variables)
  Social_Capital = c(
    "Social.Capital.Index",                # Social Capital Index
    "Fraction.Religious",                  # Fraction Religious
    "Violent.Crime.Rate"                   # Violent Crime Rate
  ),

  # FAMILY STRUCTURE (3 variables)
  Family_Structure = c(
    "Fraction.of.Children.with.Single.Mothers",  # Fraction Single Mothers
    "Fraction.of.Adults.Divorced",               # Fraction Divorced
    "Fraction.of.Adults.Married"                 # Fraction Married
  ),

  # TAX (4 variables)
  Tax = c(
    "Local.Tax.Rate",                      # Local Tax Rate
    "Local.Government.Expenditures.Per.Capita",  # Local Gov Expenditures
    "State.EITC.Exposure",                 # State EITC Exposure
    "State.Income.Tax.Progressivity"       # State Income Tax Progressivity
  ),

  # COLLEGE (3 variables)
  College = c(
    "Number.of.Colleges.per.Capita",       # Colleges per Capita
    "College.Tuition",                     # Mean College Tuition
    "College.Graduation.Rate..Income.Adjusted."  # Graduation Rate (Inc. Adj.)
  ),

  # LOCAL LABOR MARKET (4 variables)
  Local_Labor_Market = c(
    "Labor.Force.Participation.Rate",      # Labor Force Participation
    "Manufacturing.Employment.Share",      # Manufacturing Employment Share
    "Growth.in.Chinese.Imports.1990.2000", # Growth in Chinese Imports
    "Teenage.Labor.Force.Participation.Rate"  # Teen Labor Force Participation
  ),

  # MIGRATION (3 variables)
  Migration = c(
    "Migration.Inflow.Rate",               # Migration Inflow Rate
    "Migration.Outlflow.Rate",             # Migration Outflow Rate
    "Frac..Foreign.Born"                   # Fraction Foreign Born
  )
)

all_covariates <- unlist(covariates_to_test, use.names = FALSE)
cat(sprintf("Testing %d covariates across %d categories\n",
  length(all_covariates), length(covariates_to_test)))
```

# Part 4: Classical OLS Replication

## Replicating Table VIII Results

This phase replicates Chetty et al.'s classical OLS results. For each covariate, we:

1. Standardize both the covariate and relative mobility (mean 0, SD 1)
2. Regress relative mobility on the covariate
3. The standardized coefficient equals the Pearson correlation

```{r phase1}
# Initialize results dataframe
results <- data.frame(
  category = character(),
  covariate = character(),
  n_obs = integer(),
  beta = numeric(),           # Standardized coefficient (= Pearson r)
  se = numeric(),             # Standard error
  t_stat = numeric(),         # t-statistic
  p_classical = numeric(),    # Classical p-value
  stringsAsFactors = FALSE
)

# Loop through all covariates
for (category_name in names(covariates_to_test)) {
  cat(sprintf("\n%s:\n", toupper(category_name)))
  cat(paste(rep("-", 70), collapse = ""), "\n")

  for (covariate_name in covariates_to_test[[category_name]]) {

    # Check if covariate exists in data
    if (!(covariate_name %in% names(df))) {
      cat(sprintf("  %s: NOT FOUND\n", covariate_name))
      next
    }

    # Extract complete cases
    temp_data <- df %>%
      select(RM, all_of(covariate_name)) %>%
      drop_na()

    if (nrow(temp_data) < 50) {
      cat(sprintf("  %s: Insufficient data (n=%d)\n", covariate_name, nrow(temp_data)))
      next
    }

    n_obs <- nrow(temp_data)

    # Standardize variables (mean 0, SD 1)
    y <- temp_data$RM
    y_std <- (y - mean(y)) / sd(y)
    x <- temp_data[[covariate_name]]
    x_std <- (x - mean(x)) / sd(x)

    # Fit OLS regression
    fit <- lm(y_std ~ x_std)
    fit_summary <- summary(fit)
    coef_table <- coef(fit_summary)

    beta <- coef_table[2, "Estimate"]
    se <- coef_table[2, "Std. Error"]
    t_stat <- coef_table[2, "t value"]
    p_classical <- coef_table[2, "Pr(>|t|)"]

    # Store results
    results <- rbind(results, data.frame(
      category = category_name,
      covariate = covariate_name,
      n_obs = n_obs,
      beta = beta,
      se = se,
      t_stat = t_stat,
      p_classical = p_classical
    ))

    # Print result
    sig_icon <- ifelse(p_classical < 0.05, "*", " ")
    cat(sprintf("  %-35s n=%3d  b=%6.3f  p=%5.3f%s\n",
      substr(covariate_name, 1, 35), n_obs, beta, p_classical, sig_icon))
  }
}

# Summary
cat(sprintf("\n\nCLASSICAL OLS SUMMARY\n"))
cat(paste(rep("=", 70), collapse = ""), "\n")
cat(sprintf("Total covariates tested: %d\n", nrow(results)))
cat(sprintf("Significant at alpha=0.05: %d (%.1f%%)\n",
  sum(results$p_classical < 0.05),
  100 * mean(results$p_classical < 0.05)))
```

# Part 5: Permutation Test Comparison

## Comparing Classical and Permutation Inference

Now we run the permutation regression test for each covariate and compare to the classical results. This tests whether the significance conclusions are robust to the choice of inference method.

```{r phase2}
# Add permutation results columns
results$p_perm <- NA
results$classical_sig <- results$p_classical < 0.05
results$perm_sig <- NA
results$agree <- NA

# Loop through covariates
for (i in 1:nrow(results)) {
  covariate_name <- results$covariate[i]
  category_name <- results$category[i]

  if (i == 1 || results$category[i] != results$category[i - 1]) {
    cat(sprintf("\n%s:\n", toupper(category_name)))
    cat(paste(rep("-", 70), collapse = ""), "\n")
  }

  # Extract and standardize data
  temp_data <- df %>%
    select(RM, all_of(covariate_name)) %>%
    drop_na()

  y <- temp_data$RM
  y_std <- (y - mean(y)) / sd(y)
  x <- temp_data[[covariate_name]]
  x_std <- (x - mean(x)) / sd(x)

  # Center to mean zero (required for permutation test)
  x_m0 <- x_std - mean(x_std)
  y_m0 <- y_std - mean(y_std)

  # Run permutation regression test
  perm_result <- perm_test_regression(
    X = x_m0,
    Y = y_m0,
    B = 10000,
    seed = 42 + i  # Different seed for each test, but reproducible
  )

  p_perm <- perm_result$p_value
  results$p_perm[i] <- p_perm
  results$perm_sig[i] <- p_perm < 0.05
  results$agree[i] <- results$classical_sig[i] == results$perm_sig[i]

  # Print result
  agree_icon <- ifelse(results$agree[i], "[OK]", "[X]")
  sig_icon <- ifelse(results$classical_sig[i], "*", " ")

  cat(sprintf("  %s %-30s b=%6.3f  p_class=%5.3f  p_perm=%5.3f%s\n",
    agree_icon,
    substr(covariate_name, 1, 30),
    results$beta[i],
    results$p_classical[i],
    p_perm,
    sig_icon))
}
```

# Results Summary

## Agreement Between Methods

```{r summary}
cat("\n")
cat(paste(rep("=", 70), collapse = ""), "\n")
cat("RESULTS SUMMARY\n")
cat(paste(rep("=", 70), collapse = ""), "\n\n")

cat("1. CLASSICAL OLS (Replicating Chetty et al.)\n")
cat(sprintf("   - Covariates tested: %d\n", nrow(results)))
cat(sprintf("   - Significant (p < 0.05): %d (%.1f%%)\n",
  sum(results$classical_sig),
  100 * mean(results$classical_sig)))

cat("\n2. PERMUTATION REGRESSION (DiCiccio & Romano 2017)\n")
cat(sprintf("   - Significant (p < 0.05): %d (%.1f%%)\n",
  sum(results$perm_sig),
  100 * mean(results$perm_sig)))

cat("\n3. AGREEMENT ANALYSIS\n")
cat(sprintf("   - Methods agree: %d/%d (%.1f%%)\n",
  sum(results$agree),
  nrow(results),
  100 * mean(results$agree)))

# Show any disagreements
disagreements <- results %>% filter(!agree)

if (nrow(disagreements) > 0) {
  cat(sprintf("\n   DISAGREEMENTS (%d):\n", nrow(disagreements)))
  for (j in 1:nrow(disagreements)) {
    row <- disagreements[j, ]
    cat(sprintf("   - %s: Classical p=%.3f (%s), Perm p=%.3f (%s)\n",
      row$covariate,
      row$p_classical,
      ifelse(row$classical_sig, "SIG", "NOT SIG"),
      row$p_perm,
      ifelse(row$perm_sig, "SIG", "NOT SIG")))
  }
} else {
  cat("\n   No disagreements - all methods agree on significance!\n")
}

cat("\n4. P-VALUE CORRELATION\n")
cat(sprintf("   - Correlation between classical and permutation p-values: r = %.3f\n",
  cor(results$p_classical, results$p_perm)))
```

## Non-Significant Covariates

These covariates were NOT significantly associated with relative mobility under either method:

```{r nonsig}
nonsig <- results %>%
  filter(!classical_sig & !perm_sig) %>%
  select(category, covariate, beta, p_classical, p_perm)

if (nrow(nonsig) > 0) {
  cat("\nCovariates NOT significant under both methods:\n\n")
  for (j in 1:nrow(nonsig)) {
    cat(sprintf("  - %s (b=%.3f, p_class=%.3f, p_perm=%.3f)\n",
      nonsig$covariate[j], nonsig$beta[j],
      nonsig$p_classical[j], nonsig$p_perm[j]))
  }
} else {
  cat("\nAll covariates were significant under at least one method.\n")
}
```

# Save Results

```{r save-results}
# Create output directory if needed
if (!dir.exists("output")) {
  dir.create("output", recursive = TRUE)
}

# Clean up covariate names for readability
results_clean <- results %>%
  mutate(covariate_clean = case_when(
    covariate == "Frac..Black" ~ "Fraction Black Residents",
    covariate == "Racial.Segregation" ~ "Racial Segregation",
    covariate == "Income.Segregation" ~ "Income Segregation",
    covariate == "Segregation.of.Poverty...p25." ~ "Segregation of Poverty (<p25)",
    covariate == "Segregation.of.Affluence...p75." ~ "Segregation of Affluence (>p75)",
    covariate == "Frac..with.Commute...15.Mins" ~ "Share with Commute < 15 Mins",
    covariate == "Household.Income.per.capita" ~ "Household Income per Capita",
    covariate == "Gini" ~ "Gini Coefficient",
    covariate == "Top.1..Income.Share" ~ "Top 1% Income Share",
    covariate == "Gini.Bottom.99." ~ "Gini Bottom 99%",
    covariate == "Frac..Between.p25.and.p75" ~ "Fraction Middle Class",
    covariate == "School.Expenditure.per.Student" ~ "School Expenditure per Student",
    covariate == "Teacher.Student.Ratio" ~ "Teacher-Student Ratio",
    covariate == "Test.Score.Percentile..Income.Adjusted" ~ "Test Score Percentile (Inc. Adj.)",
    covariate == "High.School.Dropout.Rate..Income.Adjusted." ~ "HS Dropout Rate (Inc. Adj.)",
    covariate == "Social.Capital.Index" ~ "Social Capital Index",
    covariate == "Fraction.Religious" ~ "Fraction Religious",
    covariate == "Violent.Crime.Rate" ~ "Violent Crime Rate",
    covariate == "Fraction.of.Children.with.Single.Mothers" ~ "Fraction Single Mothers",
    covariate == "Fraction.of.Adults.Divorced" ~ "Fraction Divorced",
    covariate == "Fraction.of.Adults.Married" ~ "Fraction Married",
    covariate == "Local.Tax.Rate" ~ "Local Tax Rate",
    covariate == "Local.Government.Expenditures.Per.Capita" ~ "Local Gov Expenditures",
    covariate == "State.EITC.Exposure" ~ "State EITC Exposure",
    covariate == "State.Income.Tax.Progressivity" ~ "State Income Tax Progressivity",
    covariate == "Number.of.Colleges.per.Capita" ~ "Colleges per Capita",
    covariate == "College.Tuition" ~ "Mean College Tuition",
    covariate == "College.Graduation.Rate..Income.Adjusted." ~ "College Grad Rate (Inc. Adj.)",
    covariate == "Labor.Force.Participation.Rate" ~ "Labor Force Participation",
    covariate == "Manufacturing.Employment.Share" ~ "Manufacturing Employment Share",
    covariate == "Growth.in.Chinese.Imports.1990.2000" ~ "Growth in Chinese Imports",
    covariate == "Teenage.Labor.Force.Participation.Rate" ~ "Teen Labor Force Participation",
    covariate == "Migration.Inflow.Rate" ~ "Migration Inflow Rate",
    covariate == "Migration.Outlflow.Rate" ~ "Migration Outflow Rate",
    covariate == "Frac..Foreign.Born" ~ "Fraction Foreign Born",
    TRUE ~ gsub("\\.", " ", covariate)
  ))

# Save results
write.csv(results_clean,
  "output/mobility_test_results.csv",
  row.names = FALSE)
cat("Saved: output/mobility_test_results.csv\n")

# Save workspace
save(results, results_clean, df, covariates_to_test,
  file = "output/mobility_analysis_workspace.RData")
cat("Saved: output/mobility_analysis_workspace.RData\n")
```

# Visualization

## Coefficient Plot

This plot shows the standardized regression coefficients for all 35 covariates, with colors indicating whether both methods agree on significance.

```{r visualization, fig.width=10, fig.height=9, message=FALSE}
# Prepare data for plotting
plot_data <- results_clean %>%
  mutate(
    sig_status = case_when(
      classical_sig & perm_sig ~ "Both significant",
      !classical_sig & !perm_sig ~ "Neither significant",
      TRUE ~ "Disagreement"
    )
  ) %>%
  arrange(beta)

plot_data$covariate_clean <- factor(
  plot_data$covariate_clean,
  levels = plot_data$covariate_clean
)

# Create coefficient plot
p <- ggplot(plot_data, aes(x = beta, y = covariate_clean)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(
    aes(xmin = beta - 1.96 * se, xmax = beta + 1.96 * se, color = sig_status),
    height = 0.3, alpha = 0.7
  ) +
  geom_point(aes(color = sig_status), size = 3) +
  scale_color_manual(values = c(
    "Both significant" = "#2E8B57",
    "Neither significant" = "gray60",
    "Disagreement" = "#DC143C"
  )) +
  labs(
    title = "Predictors of Relative Mobility: Classical vs Permutation Inference",
    subtitle = "Replication of Chetty et al. (2014) Table VIII with robustness check",
    x = "Standardized Coefficient (95% CI)",
    y = "",
    color = "Significance Agreement"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.y = element_text(size = 9),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "gray40")
  )

print(p)

ggsave("output/coefficient_plot.png", p, width = 10, height = 9, dpi = 300)
cat("Saved: output/coefficient_plot.png\n")
```

# Conclusion

```{r conclusion}
cat(paste(rep("=", 70), collapse = ""), "\n")
cat("CONCLUSION\n")
cat(paste(rep("=", 70), collapse = ""), "\n\n")

agreement_pct <- 100 * mean(results$agree)

if (agreement_pct == 100) {
  cat("All 35 covariates show IDENTICAL significance conclusions under both\n")
  cat("classical OLS and permutation regression inference.\n\n")
  cat("This demonstrates that Chetty et al.'s findings are ROBUST to the\n")
  cat("choice of inference method and do not depend on asymptotic normality\n")
  cat("assumptions.\n")
} else {
  cat(sprintf("%.1f%% of covariates show identical significance conclusions.\n", agreement_pct))
  cat(sprintf("%d covariates show disagreement between methods.\n", sum(!results$agree)))
}

cat(sprintf("\n\nKey Statistics:\n"))
cat(sprintf("  - Sample size: %d Commuting Zones\n", nrow(df)))
cat(sprintf("  - Covariates tested: %d\n", nrow(results)))
cat(sprintf("  - Classically significant: %d (%.1f%%)\n",
    sum(results$classical_sig), 100*mean(results$classical_sig)))
cat(sprintf("  - Permutation significant: %d (%.1f%%)\n",
    sum(results$perm_sig), 100*mean(results$perm_sig)))
cat(sprintf("  - P-value correlation: r = %.3f\n", cor(results$p_classical, results$p_perm)))
```

# References

- Chetty, R., Hendren, N., Kline, P., & Saez, E. (2014). Where is the land of opportunity? The geography of intergenerational mobility in the United States. *Quarterly Journal of Economics*, 129(4), 1553-1623.

- DiCiccio, C. J., & Romano, J. P. (2017). Robust permutation tests for correlation and regression coefficients. *Journal of the American Statistical Association*, 112(519), 1211-1220.

- Jaimes, A. (2025). Where is the Land of Opportunity Replication. [Implementation of DiCiccio & Romano permutation tests]
